\documentclass[11pt]{article}
\begin{document}

Major knowledge gaps exist in the scientific literature regarding the mechanistic processes governing floc formation and activity. For example, it is unclear how floc formation is initiated, how enzymatic activity is regulated within flocs, what role intercellular signaling plays in floc formation and activity or why flocs lose activity as activated sludge age increases. Given how important activated sludge floc formation and activity are to human civilization it is surprising such knowledge gaps exist. A promising approach to addressing these knowledge gaps is through development and interrogation of experimental models of floc formation. The absence of such models curtails scientific endeavour in the context of wastewater treatment.  
The present study seeks to develop an experimental model system based on interactions between activated sludge bacteria and lipids. The choice of lipids as a substrate is justified by the high lipid content of wastewater and the high lipase activity in activated sludge. Further, the link between proteobacteria, which are abundant in activated sludge, and AHL dependent regulation of lipase activity offers an opportunity to explore the role of AHL mediated gene expression in activated sludge floc formation. The lipid chosen to develop the model is glycerol trioleate, which is the most abundant lipid in vegetable oil. The study is by necessity developmental with very little experimental information available as a starting point.

This project directly addresses the hypothesis that the lipid glycerol trioleate serves as a surface substratum for activated sludge floc formation. 


Specifically, the aims are:
\begin{enumerate}
\item To document the partitioning behaviour of glycerol trioleate in activated sludge.
\item To document spatial relationships between glycerol trioleate and activated sludge bacteria.
\item To describe the bacterial community composition of flocs in the presence of glycerol trioleate.
\item To isolate and identify glycerol trioleate degrading bacteria from activated sludge.
\item To test glycerol trioleate degrading bacteria from activated sludge for AHL production.
\end{enumerate}
\end{document}